### 2017-05-03 paper.tex
continuing work

\documentclass{article}

\usepackage{cmcphys}
\usepackage[margin=1.0in]{geometry}
\usepackage[titletoc,title]{appendix}
\allowdisplaybreaks

\renewcommand{\L}{\mathcal{L}}
\newcommand{\e}{\mathrm{e}}
\renewcommand{\i}{\mathrm{i}}
\renewcommand{\Re}{\mathrm{Re}\ }

\title{Chaotic Behavior of the Triple Pendulum \\
    \large A Computational Approach }
\author{Rachel Bass \and Cory McCartan}
\date{May 2017}

\begin{document}

    \maketitle

    \section{Introduction}

    \clearpage
    \begin{appendices}

    \section{Equations of Motion}
    \subsection{Full Solution}

    \begin{wrapfigure}[13]{r}{0.25\textwidth}
        \centering
        \includegraphics[width=0.2\textwidth]{drawing.tikz}
        \caption{The triple pendulum.}
    \end{wrapfigure}

    We can describe the position of a given mass by its Cartesian coordinates
    $(x_i, y_i)$.  The position of a given mass is the sum of its position
    relative to its pivot point and the position of its pivot point:
    \begin{align}
        x_1 &= l\sin\phi_1     & y_1 &= l(1-\cos\phi_1) \\
        x_2 &= x_1+l\sin\phi_2 & y_2 &= y_1 + l(1-\cos\phi_2) \\
        x_3 &= x_2+l\sin\phi_3 & y_3 &= y_2 + l(1-\cos\phi_3).
    \end{align}
    Making the appropriate substitutions, we have
    \begin{align}
        x_1 &= l\sin\phi_1     \\
        y_1 &= l(1-\cos\phi_1) \\
        x_2 &= l(\sin\phi_1+\sin\phi_2)  \\
        y_2 &= l((1-\cos\phi_1)+(1-\cos\phi_2)) \\
        x_3 &= l(\sin\phi_1+\sin\phi_2+\sin\phi_3)  \\
        y_3 &= l((1-\cos\phi_1)+(1-\cos\phi_2)+(1-\cos\phi_3)).
    \end{align}
    Taking time derivatives to find velocity, we have
    \begin{align}
        \dot x_1 &= l\dot\phi_1\cos\phi_1 \\
        \dot y_1 &= l\dot\phi_1\sin\phi_1 \\
        \dot x_2 &= l(\dot\phi_1\cos\phi_1+\dot\phi_2\cos\phi_2) \\
        \dot y_2 &= l(\dot\phi_1\sin\phi_1+\dot\phi_2\sin\phi_2) \\
        \dot x_3 &= l(\dot\phi_1\cos\phi_1+\dot\phi_2\cos\phi_2
            +\dot\phi_3\cos\phi_3) \\
        \dot y_3 &= l(\dot\phi_1\sin\phi_1+\dot\phi_2\sin\phi_2
            +\dot\phi_3\sin\phi_3).
    \end{align}
    For the kinetic energy, we will need to find $v^2={\dot x}^2{\dot y}^2$ for
    each mass. 
    \begin{align}
        v_1^2 &= {\dot x_1}^2+{\dot y_1}^2 \\
        &= (l\dot\phi_1\cos\phi_1)^2
            +(l\dot\phi_1\sin\phi_1)^2 \\
        &= l^2(\dot\phi_1^2\cos^2\phi_1+\dot\phi_1^2\sin^2\phi_1) \\
        &= l^2\dot\phi_1^2 \\\\
                %
        v_2^2 &= {\dot x_2}^2+{\dot y_2}^2 \\
        &= (l(\dot\phi_1\cos\phi_1+\dot\phi_2\cos\phi_2))^2
          +(l(\dot\phi_1\sin\phi_1+\dot\phi_2\sin\phi_2))^2 \\
        &= l^2\left( \dot\phi_1^2\cos^2\phi_1 
            + 2\dot\phi_1\dot\phi_2\cos\phi_1\cos\phi_2 
            + \dot\phi_2^2\cos^2\phi_2 + \dot\phi_1^2\sin^2\phi_1 
            + 2\dot\phi_1\dot\phi_2\sin\phi_1\sin\phi_2 
            + \dot\phi_2^2\sin^2\phi_2 \right) \\
        &= l^2\left( \dot\phi_1^2\cos^2\phi_1 + \dot\phi_1^2\sin^2\phi_1
            + \dot\phi_2^2\cos^2\phi_2 + \dot\phi_2^2\sin^2\phi_2
            + 2\dot\phi_1\dot\phi_2(\cos\phi_1\cos\phi_2 + \sin\phi_1\sin\phi_2)
             \right) \\
        &= l^2\left( \dot\phi_1^2 + \dot\phi_2^2
            + 2\dot\phi_1\dot\phi_2\cos(\phi_1 - \phi_2) \right) \\\\
                %
        v_3^2 &= {\dot x_3}^2+{\dot y_3}^2 \\
        &= (l(\dot\phi_1\cos\phi_1+\dot\phi_2\cos\phi_2+\dot\phi_3\cos\phi_3))^2 
          +(l(\dot\phi_1\sin\phi_1+\dot\phi_2\sin\phi_2+\dot\phi_3\sin\phi_3))^2 \\
        &= l^2\bigl( \dot\phi_1^2\sin^2\phi_1+\dot\phi_2^2\sin^2\phi_2
           +\dot\phi_3^2\sin^2\phi_3 + 2\dot\phi_1\dot\phi_2\sin\phi_1\sin\phi_2
           +2\dot\phi_1\dot\phi_3\sin\phi_1\sin\phi_3
           +2\dot\phi_2\dot\phi_3\sin\phi_2\sin\phi_3 \\
           &\quad +\dot\phi_1^2\cos^2\phi_1+\dot\phi_2^2\cos^2\phi_2
           +\dot\phi_3^2\cos^2\phi_3 + 2\dot\phi_1\dot\phi_2\cos\phi_1\cos\phi_2
           +2\dot\phi_1\dot\phi_3\cos\phi_1\cos\phi_3
           +2\dot\phi_2\dot\phi_3\cos\phi_2\cos\phi_3
        \bigr) \\
        &= l^2\left( \dot\phi_1^2+\dot\phi_2^2 +\dot\phi_3^2 
           +2\dot\phi_1\dot\phi_2\cos(\phi_1-\phi_2)
           +2\dot\phi_1\dot\phi_3\cos(\phi_1-\phi_3)
           +2\dot\phi_2\dot\phi_3\cos(\phi_2-\phi_3) \right) 
    \end{align}
    We can now write the kinetic and potential energy for the system by summing
    up $-mgy_i$ and $\frac{1}{2}mv_i^2$ for each mass $i$. We will drop the
    constant terms from the potential energy expression since they will be lost
    when we take the partial derivatives of the Lagrangian:
    \begin{align}
        U &= U_1 + U_2 + U_3 \\
        &= -mgl\left(3\cos\phi_1+2\cos\phi_2+\cos\phi_3 \right) \\\\
                %
        T &= T_1 + T_2 + T_3 \\
        &= \frac{1}{2}ml^2 \left( 3\dot\phi_1^2+2\dot\phi_2^2+\dot\phi_3^2
           + 4\dot\phi_1\dot\phi_2\cos(\phi_1-\phi_2)
           + 2\dot\phi_1\dot\phi_3\cos(\phi_1-\phi_3)
           + 2\dot\phi_2\dot\phi_3\cos(\phi_2-\phi_3) \right) 
    \end{align}
    So the Lagrangian is
    \begin{align}
        \L &= T + U \\
        &= -mgl\left(3\cos\phi_1+2\cos\phi_2+\cos\phi_3 \right) 
           + \frac{1}{2}ml^2 \bigl( 3\dot\phi_1^2+2\dot\phi_2^2+\dot\phi_3^2
           + 4\dot\phi_1\dot\phi_2\cos(\phi_1-\phi_2) \\
        &\qquad\qquad + 2\dot\phi_1\dot\phi_3\cos(\phi_1-\phi_3)
           + 2\dot\phi_2\dot\phi_3\cos(\phi_2-\phi_3) \bigr) 
    \end{align}
    Since our generalized coordinates are $\phi_1,\phi_2$, and $\phi_3$, we must
    now take partial derivatives of the Lagrangian with respect to each $\phi_i$
    and $\dot\phi_i$:
    \begin{align}
        \pD{\L}{\phi_1} &= 3mgl\sin\phi_1 
            -2ml^2\dot\phi_1\dot\phi_2\sin(\phi_1-\phi_2)
            -ml^2\dot\phi_1\dot\phi_3\sin(\phi_1-\phi_3) \\
        \pD{\L}{\phi_2} &= 2mgl\sin\phi_2 
            +2ml^2\dot\phi_1\dot\phi_2\sin(\phi_1-\phi_2)
            -ml^2\dot\phi_2\dot\phi_3\sin(\phi_2-\phi_3) \\
        \pD{\L}{\phi_3} &= mgl\sin\phi_3 
            +ml^2\dot\phi_1\dot\phi_3\sin(\phi_1-\phi_3)
            +ml^2\dot\phi_2\dot\phi_3\sin(\phi_2-\phi_3) \\\\
                %
        \pD{\L}{\dot\phi_1} &= 3ml^2\dot\phi_1 
            + 2ml^2\dot\phi_2\cos(\phi_1-\phi_2) 
            + ml^2\dot\phi_3\cos(\phi_1-\phi_3)  \\
        \pD{\L}{\dot\phi_2} &= 2ml^2\dot\phi_2 
            + 2ml^2\dot\phi_1\cos(\phi_1-\phi_2) 
            + ml^2\dot\phi_3\cos(\phi_2-\phi_3)  \\
        \pD{\L}{\dot\phi_3} &= ml^2\dot\phi_3 
            + ml^2\dot\phi_1\cos(\phi_1-\phi_3) 
            + ml^2\dot\phi_2\cos(\phi_2-\phi_3) 
    \end{align}
    Next, we find $\D{}{t}\pD{\L}{\dot\phi_i}$ for each $i$:
    \begin{align}
        \D{}{t}\pD{\L}{\dot\phi_1} &= 3ml^2\ddot\phi_1 
            + 2ml^2\ddot\phi_2\cos(\phi_1-\phi_2) 
            - 2ml^2\dot\phi_2\sin(\phi_1-\phi_2)(\dot\phi_1-\dot\phi_2) \\
            &\quad + ml^2\ddot\phi_3\cos(\phi_1-\phi_3) 
            - ml^2\dot\phi_3\sin(\phi_1-\phi_3)(\dot\phi_1-\dot\phi_3) \\\\
        \D{}{t}\pD{\L}{\dot\phi_2} &= 2ml^2\ddot\phi_2 
            + 2ml^2\ddot\phi_1\cos(\phi_1-\phi_2) 
            - 2ml^2\dot\phi_1\sin(\phi_1-\phi_2)(\dot\phi_1-\dot\phi_2) \\
            &\quad + ml^2\ddot\phi_3\cos(\phi_2-\phi_3) 
            - ml^2\dot\phi_3\sin(\phi_2-\phi_3)(\dot\phi_2-\dot\phi_3) \\\\
        \D{}{t}\pD{\L}{\dot\phi_3} &= ml^2\ddot\phi_3 
            + ml^2\ddot\phi_1\cos(\phi_1-\phi_3) 
            - ml^2\dot\phi_1\sin(\phi_1-\phi_3)(\dot\phi_1-\dot\phi_3) \\
            &\quad + ml^2\ddot\phi_2\cos(\phi_2-\phi_3) 
            - ml^2\dot\phi_2\sin(\phi_2-\phi_3)(\dot\phi_2-\dot\phi_3) 
    \end{align}
    We can then write the equations of motion for each of our generalized
    coordinates, according to the Euler-Lagrange condition $\pD{\L}{\phi_i}
    =\D{}{t}\pD{\L}{\dot\phi_i}$. For $\phi_1$ we have
    \begin{align}
        \pD{\L}{\phi_1} &= \D{}{t}\pD{\L}{\dot\phi_1} 
    \end{align}
    \begin{align}
        3mgl\sin\phi_1 -2ml^2\dot\phi_1\dot\phi_2\sin(\phi_1-\phi_2)
            -ml^2\dot\phi_1\dot\phi_3\sin(\phi_1-\phi_3)
        &= 3ml^2\ddot\phi_1 + 2ml^2\ddot\phi_2\cos(\phi_1-\phi_2) \\
        &\quad - 2ml^2\dot\phi_2\sin(\phi_1-\phi_2)(\dot\phi_1-\dot\phi_2) \\
        &\quad + ml^2\ddot\phi_3\cos(\phi_1-\phi_3) \\
        &\quad - ml^2\dot\phi_3\sin(\phi_1-\phi_3)(\dot\phi_1-\dot\phi_3) 
    \end{align}
    \begin{align}
        3mgl\sin\phi_1 &= 3ml^2\ddot\phi_1 + 2ml^2\ddot\phi_2\cos(\phi_1-\phi_2)
            + ml^2\ddot\phi_3\cos(\phi_1-\phi_3)  \\
            &\quad + 2ml^2\dot\phi_2^2\sin(\phi_1-\phi_2)
            - ml^2\dot\phi_3^2\sin(\phi_1-\phi_3)
    \end{align}
    \begin{align}
        \frac{3g}{l}\sin\phi_1 &= 3\ddot\phi_1 + 2\ddot\phi_2\cos(\phi_1-\phi_2)
            + \ddot\phi_3\cos(\phi_1-\phi_3) + 2\dot\phi_2^2\sin(\phi_1-\phi_2)
            - \dot\phi_3^2\sin(\phi_1-\phi_3)   \label{eq_m1}
    \end{align}

    For $\phi_2$ we have
    \begin{align}
        \pD{\L}{\phi_2} &= \D{}{t}\pD{\L}{\dot\phi_2} 
    \end{align}
    \begin{align}
        2mgl\sin\phi_2 +2ml^2\dot\phi_1\dot\phi_2\sin(\phi_1-\phi_2)
            -ml^2\dot\phi_2\dot\phi_3\sin(\phi_2-\phi_3)
        &= 2ml^2\ddot\phi_2 + 2ml^2\ddot\phi_1\cos(\phi_1-\phi_2) \\
        &\quad - 2ml^2\dot\phi_1\sin(\phi_1-\phi_2)(\dot\phi_1-\dot\phi_2) \\
        &\quad + ml^2\ddot\phi_3\cos(\phi_2-\phi_3) \\
        &\quad - ml^2\dot\phi_3\sin(\phi_2-\phi_3)(\dot\phi_2-\dot\phi_3) 
    \end{align}
    \begin{align}
        2mgl\sin\phi_2 &= 2ml^2\ddot\phi_2 + 2ml^2\ddot\phi_1\cos(\phi_1-\phi_2)
            + ml^2\ddot\phi_3\cos(\phi_2-\phi_3)  \\
            &\quad + 2ml^2\dot\phi_1^2\sin(\phi_1-\phi_2)
            - ml^2\dot\phi_3^2\sin(\phi_2-\phi_3)
    \end{align}
    \begin{align}
        \frac{2g}{l}\sin\phi_2 &= 2\ddot\phi_2 + 2\ddot\phi_1\cos(\phi_1-\phi_2)
            + \ddot\phi_3\cos(\phi_2-\phi_3) + 2\dot\phi_1^2\sin(\phi_1-\phi_2)
            - \dot\phi_3^2\sin(\phi_2-\phi_3)   \label{eq_m2}
    \end{align}

    For $\phi_3$ we have
    \begin{align}
        \pD{\L}{\phi_3} &= \D{}{t}\pD{\L}{\dot\phi_3} 
    \end{align}
    \begin{align}
        mgl\sin\phi_3 +ml^2\dot\phi_1\dot\phi_3\sin(\phi_1-\phi_3)
            +ml^2\dot\phi_2\dot\phi_3\sin(\phi_2-\phi_3)
        &= ml^2\ddot\phi_3 + ml^2\ddot\phi_1\cos(\phi_1-\phi_3) \\
        &\quad - ml^2\dot\phi_1\sin(\phi_1-\phi_3)(\dot\phi_1-\dot\phi_3) \\
        &\quad + ml^2\ddot\phi_2\cos(\phi_2-\phi_3) \\
        &\quad - ml^2\dot\phi_2\sin(\phi_2-\phi_3)(\dot\phi_2-\dot\phi_3) 
    \end{align}
    \begin{align}
        mgl\sin\phi_3 &= ml^2\ddot\phi_3 + ml^2\ddot\phi_1\cos(\phi_1-\phi_3)
            + ml^2\ddot\phi_2\cos(\phi_2-\phi_3)  \\
            &\quad + ml^2\dot\phi_1^2\sin(\phi_1-\phi_3)
            - ml^2\dot\phi_2^2\sin(\phi_2-\phi_3)
    \end{align}
    \begin{align}
        \frac{g}{l}\sin\phi_3 &= \ddot\phi_3 + \ddot\phi_1\cos(\phi_1-\phi_3)
            + \ddot\phi_2\cos(\phi_2-\phi_3) + \dot\phi_1^2\sin(\phi_1-\phi_3)
            - \dot\phi_2^2\sin(\phi_2-\phi_3)   \label{eq_m3}
    \end{align}

    Rearranging $\eqref{eq_m1}$, $\eqref{eq_m2}$, and $\eqref{eq_m3}$ by moving
    the second derivatives to one side, we have
    \begin{align}
         3\ddot\phi_1 + 2\ddot\phi_2\cos(\phi_1-\phi_2) + \ddot\phi_3\cos(\phi_1-\phi_3) + 
             &= -2\dot\phi_2^2\sin(\phi_1-\phi_2) + \dot\phi_3^2\sin(\phi_1-\phi_3)
             +\frac{3g}{l}\sin\phi_1 \\
        2\ddot\phi_1\cos(\phi_1-\phi_2)+ 2\ddot\phi_2 + \ddot\phi_3\cos(\phi_2-\phi_3)
            &= - 2\dot\phi_1^2\sin(\phi_1-\phi_2) + \dot\phi_3^2\sin(\phi_2-\phi_3)
            +\frac{2g}{l}\sin\phi_2 \\
        \ddot\phi_1\cos(\phi_1-\phi_3+ \ddot\phi_2\cos(\phi_2-\phi_3)) + \ddot\phi_3
            &= - \dot\phi_1^2\sin(\phi_1-\phi_3) + \dot\phi_2^2\sin(\phi_2-\phi_3)
            +\frac{g}{l}\sin\phi_3
    \end{align}
    We can rewrite this system  equations as a single matrix equation:
    \begingroup 
    \begin{equation} 
        \renewcommand*{\arraystretch}{1.75} % more vertical space in matrix
        \resizebox{0.95\textwidth}{!}{$
        \begin{pmatrix}
            3                    & 2\cos(\phi_1-\phi_2) & \cos(\phi_1-\phi_3) \\
            2\cos(\phi_1-\phi_2) & 2                    & \cos(\phi_2-\phi_3) \\
            \cos(\phi_1-\phi_3)  & \cos(\phi_2-\phi_3)  & 1                   \\
        \end{pmatrix}
        \begin{pmatrix}
            \ddot\phi_1 \\
            \ddot\phi_2 \\
            \ddot\phi_3
        \end{pmatrix} =
        \begin{pmatrix}
            \dot\phi_3^2\sin(\phi_1-\phi_3) -2\dot\phi_2^2\sin(\phi_1-\phi_2)
                +\frac{3g}{l}\sin\phi_1 \\
            \dot\phi_3^2\sin(\phi_2-\phi_3) - 2\dot\phi_1^2\sin(\phi_1-\phi_2)
                +\frac{2g}{l}\sin\phi_2 \\
            \dot\phi_2^2\sin(\phi_2-\phi_3) -\dot\phi_1^2\sin(\phi_1-\phi_3)
                +\frac{g}{l}\sin\phi_3
        \end{pmatrix}
        $}
    \end{equation} 
    \endgroup 
    We can then solve for our vector of second derivatives by multiplying on the
    left by the inverse of the first matrix:
    \begingroup 
    \begin{equation} \renewcommand*{\arraystretch}{1.75} % more vertical space in matrix
        \resizebox{0.95\textwidth}{!}{$
        \begin{pmatrix}
            \ddot\phi_1 \\
            \ddot\phi_2 \\
            \ddot\phi_3
        \end{pmatrix} =
        \begin{pmatrix}
            3                    & 2\cos(\phi_1-\phi_2) & \cos(\phi_1-\phi_3) \\
            2\cos(\phi_1-\phi_2) & 2                    & \cos(\phi_2-\phi_3) \\
            \cos(\phi_1-\phi_3)  & \cos(\phi_2-\phi_3)  & 1                   \\
        \end{pmatrix}^{-1}
        \begin{pmatrix}
            \dot\phi_3^2\sin(\phi_1-\phi_3) -2\dot\phi_2^2\sin(\phi_1-\phi_2)
                +\frac{3g}{l}\sin\phi_1 \\
            \dot\phi_3^2\sin(\phi_2-\phi_3) - 2\dot\phi_1^2\sin(\phi_1-\phi_2)
                +\frac{2g}{l}\sin\phi_2 \\
            \dot\phi_2^2\sin(\phi_2-\phi_3) -\dot\phi_1^2\sin(\phi_1-\phi_3)
                +\frac{g}{l}\sin\phi_3
        \end{pmatrix} 
        $} \label{mat-eqn}
    \end{equation} 
    \endgroup 
    The matrix inverse in the above equation is very tedious to calculate. We
    used a computer algebra system\footnote{SageMath, the Sage Mathematics
        Software System (Version 7.6), The Sage Developers, 2017,
        \texttt{http://www.sagemath.org}.}
    to do the inverse and the matrix multiplication afterwards. The result is
    \begingroup 
    \begin{equation} 
        \renewcommand*{\arraystretch}{2.5} % more vertical space in matrix
        \resizebox{0.98\textwidth}{!}{$
        \begin{pmatrix}
            \ddot\phi_1 \\
            \ddot\phi_2 \\
            \ddot\phi_3
        \end{pmatrix} =
        \begin{pmatrix}
            \frac{1}{3} {\left(2 \dot\phi_2^2 \sin(\phi_1-\phi_2)-\dot\phi_3^2 \sin(\phi_1-\phi_3)-\frac{3 g \sin\left( \phi_1\right)}{l}\right)} {\left(\frac{2 {\left(\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)} \cos(\phi_1-\phi_2)}{2 \cos(\phi_1-\phi_2)^2- 3}-\cos(\phi_1-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6} + \frac{2 \cos(\phi_1-\phi_2)^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 1\right)}- {\left(2 \dot\phi_1^2 \sin(\phi_1-\phi_2)-\dot\phi_3^2 \sin(\phi_2-\phi_3)-\frac{2 g \sin\left(\phi_1\right)}{l}\right)} {\left(\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)} {\left(\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)} \cos(\phi_1-\phi_2)}{2 \cos(\phi_1-\phi_2)^2- 3}-\cos(\phi_1-\phi_3)\right)}}{{\left(2 \cos(\phi_1-\phi_2)^2- 3\right)} {\left(2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6\right)}} + \frac{\cos(\phi_1-\phi_2)}{2 \cos(\phi_1-\phi_2)^2- 3}\right)} + \frac{2 {\left(\dot\phi_1^2 \sin(\phi_1-\phi_3)-\dot\phi_2^2 \sin(\phi_2-\phi_3)-\frac{g \sin\left(\phi_1\right)}{l}\right)} {\left(\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)} \cos(\phi_1-\phi_2)}{2 \cos(\phi_1-\phi_2)^2- 3}-\cos(\phi_1-\phi_3)\right)}}{2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6} \\
            -{\left(2 \dot\phi_2^2 \sin(\phi_1-\phi_2)-\dot\phi_3^2 \sin(\phi_1-\phi_3)-\frac{3 g\sin\left(\phi_1\right)}{l}\right)} {\left(\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)} {\left(\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)} \cos(\phi_1-\phi_2)}{2 \cos(\phi_1-\phi_2)^2- 3}-\cos(\phi_1-\phi_3)\right)}}{{\left(2 \cos(\phi_1-\phi_2)^2- 3\right)} {\left(2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6\right)}} + \frac{\cos(\phi_1-\phi_2)}{2 \cos(\phi_1-\phi_2)^2- 3}\right)} + \frac{3}{2} {\left(2 \dot\phi_1^2 \sin(\phi_1-\phi_2)-\dot\phi_3^2 \sin(\phi_2-\phi_3)-\frac{2 g \sin\left(\phi_1\right)}{l}\right)} {\left(\frac{1}{2 \cos(\phi_1-\phi_2)^2- 3} + \frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{{\left(2 \cos(\phi_1-\phi_2)^2- 3\right)}^2 {\left(2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6\right)}}\right)}-\frac{3 {\left(\dot\phi_1^2 \sin(\phi_1-\phi_3)-\dot\phi_2^2 \sin(\phi_2-\phi_3)-\frac{g \sin\left(\phi_1\right)}{l}\right)} {\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}}{{\left(2 \cos(\phi_1-\phi_2)^2- 3\right)} {\left(2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6\right)}} \\
           \frac{2 {\left(2 \dot\phi_2^2 \sin(\phi_1-\phi_2)-\dot\phi_3^2 \sin(\phi_1-\phi_3)-\frac{3 g \sin\left(\phi_1\right)}{l}\right)} {\left(\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)} \cos(\phi_1-\phi_2)}{2 \cos(\phi_1-\phi_2)^2- 3}-\cos(\phi_1-\phi_3)\right)}}{2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6} + \frac{6 {\left(\dot\phi_1^2 \sin(\phi_1-\phi_3)-\dot\phi_2^2 \sin(\phi_2-\phi_3)-\frac{g \sin\left(\phi_1\right)}{l}\right)}}{2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6}-\frac{3 {\left(2 \dot\phi_1^2 \sin(\phi_1-\phi_2)-\dot\phi_3^2 \sin(\phi_2-\phi_3)-\frac{2 g \sin\left(\phi_1\right)}{l}\right)} {\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}}{{\left(2 \cos(\phi_1-\phi_2)^2- 3\right)} {\left(2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6\right)}}
        \end{pmatrix}
        $}\, ,\vspace{0.5em}
    \end{equation} 
    \endgroup 
    which is far too messy to use computationally. Instead, we chose to
    construct the matrices in $\eqref{mat-eqn}$ in \texttt{numpy} and
    numerically calculate the inverse and multiplication. Since many of the
    terms are repeated (especially the trigonometric functions of angle
    differences), this has the added advantage of allowing us to precompute
    these functions.

    \subsection{Small-angle Approximation}

    To check the validity of our numerical solution, we would like an analytical
    solution. Clearly a full analytical solution is impossible. However, by
    assuming that $\phi_1,\phi_2$, and $\phi_3$ remain small, we can radically
    simplify the equations of motion to reach a closed-form expression for
    $\phi_1,\phi_2$, and $phi_3$ as a function of time.

    We begin by substituting $\cos\phi_i=1-\frac{1}{2}\phi_i^2$, the second
    order Maclaurin approximation for cosine, into our expression for potential
    energy. We also substitute $\dot\phi_i\dot\phi_j\cos(\phi_i-\phi_j)=\dot
    \phi_i\dot\phi_j$, since for the small product $\dot\phi_i\dot\phi_j$, the
    cosine term is negligible. This allows us to simplify our expression for
    kinetic energy as well. We find
    \begin{align}
        U &= -mgl\left(3(1-\tfrac{1}{2}\phi_1^2)+2(1-\tfrac{1}{2}\phi_2^2
            +(1-\tfrac{1}{2}\phi_3^2))\right) \\
        &\simeq mgl\left(\frac{3}{2}\phi_1^2+\frac{2}{2}\phi_2^2
            +\frac{1}{2}\phi_3^2\right) \tag*{(since the constant terms will be
            dropped when the derivative is taken)} \\\\
        T &= \frac{1}{2}ml^2(3\dot\phi_1^2+2\dot\phi_2^2+\dot\phi_1^2
            +4\dot\phi_1\phi_2+2\dot\phi_1\phi_3+2\dot\phi_2\phi_3),
    \end{align}
    so the Lagrangian is
    \begin{align}
        \L &= T+U \\
        &= \frac{1}{2}ml^2(3\dot\phi_1^2+2\dot\phi_2^2+\dot\phi_1^2
            +4\dot\phi_1\phi_2+2\dot\phi_1\phi_3+2\dot\phi_2\phi_3)
            -mgl\left(\frac{3}{2}\phi_1^2+\frac{2}{2}\phi_2^2
            +\frac{1}{2}\phi_3^2\right)
    \end{align}
    Taking derivatives as required, we have
    \begin{align}
        \pD{\L}{\phi_1} &= -3mgl\phi_1, & 
        \pD{\L}{\dot\phi_1} &= ml^2(3\dot\phi_1+4\dot\phi_2+2\dot\phi_3)
        &\Rightarrow\quad
        \D{}{t}\pD{\L}{\dot\phi_1} &= ml^2(3\ddot\phi_1+4\ddot\phi_2+2\ddot\phi_3) \\
                %
        \pD{\L}{\phi_2} &= -2mgl\phi_2, & 
        \pD{\L}{\dot\phi_2} &= ml^2(4\dot\phi_1+2\dot\phi_2+2\dot\phi_3)
        &\Rightarrow\quad
        \D{}{t}\pD{\L}{\dot\phi_3} &= ml^2(4\ddot\phi_1+2\ddot\phi_2+2\ddot\phi_3) \\
                %
        \pD{\L}{\phi_3} &= -mgl\phi_3, & 
        \pD{\L}{\dot\phi_3} &= ml^2(2\dot\phi_1+2\dot\phi_2+\dot\phi_3)
        &\Rightarrow\quad
        \D{}{t}\pD{\L}{\dot\phi_3} &= ml^2(2\ddot\phi_1+2\ddot\phi_2+\ddot\phi_3) \\
    \end{align}
    Applying our equilibrium condition $\D{}{t}\pD{\L}{\dot\phi_i}=\pD{\L}
    {\phi_i}$, we have our three equations of motion:
    \begin{align}
        ml^2(3\ddot\phi_1+4\ddot\phi_2+2\ddot\phi_3) &= -3mgl\phi_1 \\
        ml^2(4\ddot\phi_1+2\ddot\phi_2+2\ddot\phi_3) &= -2mgl\phi_2 \\
        ml^2(2\ddot\phi_1+2\ddot\phi_2+\ddot\phi_3) &= -mgl\phi_3
    \end{align}
    Cancelling the common term of $ml^2$, we have
    \begin{align}
        3\ddot\phi_1+4\ddot\phi_2+2\ddot\phi_3 &= -\frac{3g}{l}\phi_1 \\
        4\ddot\phi_1+2\ddot\phi_2+2\ddot\phi_3 &= -\frac{2g}{l}\phi_2 \\
        2\ddot\phi_1+2\ddot\phi_2+\ddot\phi_3  &= -\frac{g}{l}\phi_3
    \end{align}
    We can write this system of equations as a single matrix equation \[
        M\vec{\ddot\phi} = -G\vec{\phi} 
    \] where
    \begin{align}
        M &= \begin{pmatrix}3&4&2\\4&2&2\\2&2&1\end{pmatrix} &
        G &= \frac{g}{l}\begin{pmatrix}3&0&0\\0&2&0\\0&0&1\end{pmatrix}
    \end{align}
    We assume a periodic solution of the form \[
        \vec\phi = \Re \vec{z}, \quad\mathrm{where}\quad
        \vec{z} = \vec{a}\,\e^{\i\omega t}
    \] Taking the second derivative of $\vec{z}$, we find \[
        \vec{\ddot z} = -\omega^2\vec{a}\,\e^{\i\omega t} = -omega^2\vec{z}
    \] So our differential equation becomes \[
        -\omega^2 M\vec{a}\e^{\i\omega t} = -G\vec{a}\e^{\i\omega t},
    \] which by rearranging and cancelling the common exponential term yields \[
        (G-\omega^2M)\vec{a} = 0
    \] This equation will have nontrivial solutions when $\det(G-\omega^2M)=0$.
    Calculating the determinant (again with a computer algebra system), we find
    \begin{align}
        w^{6}+\frac{9g}{l}w^4+\frac{9g^2}{l^2}w^2-\frac{3g^3}{l^3} &= 0
    \end{align}

    \end{appendices}

\end{document}