### 2017-05-09 paper.tex
further work

\documentclass{article}

\usepackage{cmcphys}
\usepackage[margin=1.0in]{geometry}
\usepackage[titletoc,title]{appendix}
\allowdisplaybreaks

\renewcommand{\L}{\mathcal{L}}
\newcommand{\e}{\mathrm{e}}
\renewcommand{\i}{\mathrm{i}}
\renewcommand{\Re}{\mathrm{Re}\ }

\title{Chaotic Behavior of the Triple Pendulum \\
    \large A Computational Approach }
\author{Rachel Bass \and Cory McCartan}
\date{May 2017}

\begin{document}

    \maketitle

    \section{Introduction}

    \clearpage
    \begin{appendices}

    \section{Equations of Motion}

    \begin{wrapfigure}[13]{r}{0.25\textwidth}
        \centering
        \includegraphics[width=0.2\textwidth]{drawing.tikz}
        \caption{The triple pendulum.}
    \end{wrapfigure}

    We can describe the position of a given mass by its Cartesian coordinates
    $(x_i, y_i)$.  The position of a given mass is the sum of its position
    relative to its pivot point and the position of its pivot point:
    \begin{align}
        x_1 &= l\sin\phi_1     & y_1 &= -l\cos\phi_1 \\
        x_2 &= x_1+l\sin\phi_2 & y_2 &= y_1 + -l\cos\phi_2 \\
        x_3 &= x_2+l\sin\phi_3 & y_3 &= y_2 + -l\cos\phi_3.
    \end{align}
    Making the appropriate substitutions, we have
    \begin{align}
        x_1 &= l\sin\phi_1     \\
        y_1 &= -l\cos\phi_1 \\
        x_2 &= l(\sin\phi_1+\sin\phi_2)  \\
        y_2 &= -l(\cos\phi_1+\cos\phi_2) \\
        x_3 &= l(\sin\phi_1+\sin\phi_2+\sin\phi_3)  \\
        y_3 &= -l(\cos\phi_1+\cos\phi_2+\cos\phi_3).
    \end{align}
    Taking time derivatives to find velocity, we have
    \begin{align}
        \dot x_1 &= l\dot\phi_1\cos\phi_1 \\
        \dot y_1 &= l\dot\phi_1\sin\phi_1 \\
        \dot x_2 &= l(\dot\phi_1\cos\phi_1+\dot\phi_2\cos\phi_2) \\
        \dot y_2 &= l(\dot\phi_1\sin\phi_1+\dot\phi_2\sin\phi_2) \\
        \dot x_3 &= l(\dot\phi_1\cos\phi_1+\dot\phi_2\cos\phi_2
            +\dot\phi_3\cos\phi_3) \\
        \dot y_3 &= l(\dot\phi_1\sin\phi_1+\dot\phi_2\sin\phi_2
            +\dot\phi_3\sin\phi_3).
    \end{align}
    For the kinetic energy, we will need to find $v^2={\dot x}^2{\dot y}^2$ for
    each mass. 
    \begin{align}
        v_1^2 &= {\dot x_1}^2+{\dot y_1}^2 \\
        &= (l\dot\phi_1\cos\phi_1)^2
            +(l\dot\phi_1\sin\phi_1)^2 \\
        &= l^2(\dot\phi_1^2\cos^2\phi_1+\dot\phi_1^2\sin^2\phi_1) \\
        &= l^2\dot\phi_1^2 \\\\
                %
        v_2^2 &= {\dot x_2}^2+{\dot y_2}^2 \\
        &= (l(\dot\phi_1\cos\phi_1+\dot\phi_2\cos\phi_2))^2
          +(l(\dot\phi_1\sin\phi_1+\dot\phi_2\sin\phi_2))^2 \\
        &= l^2\left( \dot\phi_1^2\cos^2\phi_1 
            + 2\dot\phi_1\dot\phi_2\cos\phi_1\cos\phi_2 
            + \dot\phi_2^2\cos^2\phi_2 + \dot\phi_1^2\sin^2\phi_1 
            + 2\dot\phi_1\dot\phi_2\sin\phi_1\sin\phi_2 
            + \dot\phi_2^2\sin^2\phi_2 \right) \\
        &= l^2\left( \dot\phi_1^2\cos^2\phi_1 + \dot\phi_1^2\sin^2\phi_1
            + \dot\phi_2^2\cos^2\phi_2 + \dot\phi_2^2\sin^2\phi_2
            + 2\dot\phi_1\dot\phi_2(\cos\phi_1\cos\phi_2 + \sin\phi_1\sin\phi_2)
             \right) \\
        &= l^2\left( \dot\phi_1^2 + \dot\phi_2^2
            + 2\dot\phi_1\dot\phi_2\cos(\phi_1 - \phi_2) \right) \\\\
                %
        v_3^2 &= {\dot x_3}^2+{\dot y_3}^2 \\
        &= (l(\dot\phi_1\cos\phi_1+\dot\phi_2\cos\phi_2+\dot\phi_3\cos\phi_3))^2 
          +(l(\dot\phi_1\sin\phi_1+\dot\phi_2\sin\phi_2+\dot\phi_3\sin\phi_3))^2 \\
        &= l^2\bigl( \dot\phi_1^2\sin^2\phi_1+\dot\phi_2^2\sin^2\phi_2
           +\dot\phi_3^2\sin^2\phi_3 + 2\dot\phi_1\dot\phi_2\sin\phi_1\sin\phi_2
           +2\dot\phi_1\dot\phi_3\sin\phi_1\sin\phi_3
           +2\dot\phi_2\dot\phi_3\sin\phi_2\sin\phi_3 \\
           &\quad +\dot\phi_1^2\cos^2\phi_1+\dot\phi_2^2\cos^2\phi_2
           +\dot\phi_3^2\cos^2\phi_3 + 2\dot\phi_1\dot\phi_2\cos\phi_1\cos\phi_2
           +2\dot\phi_1\dot\phi_3\cos\phi_1\cos\phi_3
           +2\dot\phi_2\dot\phi_3\cos\phi_2\cos\phi_3
        \bigr) \\
        &= l^2\left( \dot\phi_1^2+\dot\phi_2^2 +\dot\phi_3^2 
           +2\dot\phi_1\dot\phi_2\cos(\phi_1-\phi_2)
           +2\dot\phi_1\dot\phi_3\cos(\phi_1-\phi_3)
           +2\dot\phi_2\dot\phi_3\cos(\phi_2-\phi_3) \right) 
    \end{align}
    We can now write the kinetic and potential energy for the system by summing
    up $-mgy_i$ and $\frac{1}{2}mv_i^2$ for each mass $i$:
    \begin{align}
        U &= U_1 + U_2 + U_3 \\
        &= -mgl\left(3\cos\phi_1+2\cos\phi_2+\cos\phi_3 \right) \\\\
                %
        T &= T_1 + T_2 + T_3 \\
        &= \frac{1}{2}ml^2 \left( 3\dot\phi_1^2+2\dot\phi_2^2+\dot\phi_3^2
           + 4\dot\phi_1\dot\phi_2\cos(\phi_1-\phi_2)
           + 2\dot\phi_1\dot\phi_3\cos(\phi_1-\phi_3)
           + 2\dot\phi_2\dot\phi_3\cos(\phi_2-\phi_3) \right) 
    \end{align}
    So the Lagrangian is
    \begin{align}
        \L &= T - U \\
        &= mgl\left(3\cos\phi_1+2\cos\phi_2+\cos\phi_3 \right) 
           + \frac{1}{2}ml^2 \bigl( 3\dot\phi_1^2+2\dot\phi_2^2+\dot\phi_3^2
           + 4\dot\phi_1\dot\phi_2\cos(\phi_1-\phi_2) \\
        &\qquad\qquad + 2\dot\phi_1\dot\phi_3\cos(\phi_1-\phi_3)
           + 2\dot\phi_2\dot\phi_3\cos(\phi_2-\phi_3) \bigr) 
    \end{align}
    Since our generalized coordinates are $\phi_1,\phi_2$, and $\phi_3$, we must
    now take partial derivatives of the Lagrangian with respect to each $\phi_i$
    and $\dot\phi_i$:
    \begin{align}
        \pD{\L}{\phi_1} &= -3mgl\sin\phi_1 
            -2ml^2\dot\phi_1\dot\phi_2\sin(\phi_1-\phi_2)
            -ml^2\dot\phi_1\dot\phi_3\sin(\phi_1-\phi_3) \\
        \pD{\L}{\phi_2} &= -2mgl\sin\phi_2 
            +2ml^2\dot\phi_1\dot\phi_2\sin(\phi_1-\phi_2)
            -ml^2\dot\phi_2\dot\phi_3\sin(\phi_2-\phi_3) \\
        \pD{\L}{\phi_3} &= -mgl\sin\phi_3 
            +ml^2\dot\phi_1\dot\phi_3\sin(\phi_1-\phi_3)
            +ml^2\dot\phi_2\dot\phi_3\sin(\phi_2-\phi_3) \\\\
                %
        \pD{\L}{\dot\phi_1} &= 3ml^2\dot\phi_1 
            + 2ml^2\dot\phi_2\cos(\phi_1-\phi_2) 
            + ml^2\dot\phi_3\cos(\phi_1-\phi_3)  \\
        \pD{\L}{\dot\phi_2} &= 2ml^2\dot\phi_2 
            + 2ml^2\dot\phi_1\cos(\phi_1-\phi_2) 
            + ml^2\dot\phi_3\cos(\phi_2-\phi_3)  \\
        \pD{\L}{\dot\phi_3} &= ml^2\dot\phi_3 
            + ml^2\dot\phi_1\cos(\phi_1-\phi_3) 
            + ml^2\dot\phi_2\cos(\phi_2-\phi_3) 
    \end{align}
    Next, we find $\D{}{t}\pD{\L}{\dot\phi_i}$ for each $i$:
    \begin{align}
        \D{}{t}\pD{\L}{\dot\phi_1} &= 3ml^2\ddot\phi_1 
            + 2ml^2\ddot\phi_2\cos(\phi_1-\phi_2) 
            - 2ml^2\dot\phi_2\sin(\phi_1-\phi_2)(\dot\phi_1-\dot\phi_2) \\
            &\quad + ml^2\ddot\phi_3\cos(\phi_1-\phi_3) 
            - ml^2\dot\phi_3\sin(\phi_1-\phi_3)(\dot\phi_1-\dot\phi_3) \\\\
        \D{}{t}\pD{\L}{\dot\phi_2} &= 2ml^2\ddot\phi_2 
            + 2ml^2\ddot\phi_1\cos(\phi_1-\phi_2) 
            - 2ml^2\dot\phi_1\sin(\phi_1-\phi_2)(\dot\phi_1-\dot\phi_2) \\
            &\quad + ml^2\ddot\phi_3\cos(\phi_2-\phi_3) 
            - ml^2\dot\phi_3\sin(\phi_2-\phi_3)(\dot\phi_2-\dot\phi_3) \\\\
        \D{}{t}\pD{\L}{\dot\phi_3} &= ml^2\ddot\phi_3 
            + ml^2\ddot\phi_1\cos(\phi_1-\phi_3) 
            - ml^2\dot\phi_1\sin(\phi_1-\phi_3)(\dot\phi_1-\dot\phi_3) \\
            &\quad + ml^2\ddot\phi_2\cos(\phi_2-\phi_3) 
            - ml^2\dot\phi_2\sin(\phi_2-\phi_3)(\dot\phi_2-\dot\phi_3) 
    \end{align}
    We can then write the equations of motion for each of our generalized
    coordinates, according to the Euler-Lagrange condition $\pD{\L}{\phi_i}
    =\D{}{t}\pD{\L}{\dot\phi_i}$. For $\phi_1$ we have
    \begin{align}
        \pD{\L}{\phi_1} &= \D{}{t}\pD{\L}{\dot\phi_1} 
    \end{align}
    \begin{align}
        -3mgl\sin\phi_1 -2ml^2\dot\phi_1\dot\phi_2\sin(\phi_1-\phi_2)
            -ml^2\dot\phi_1\dot\phi_3\sin(\phi_1-\phi_3)
        &= 3ml^2\ddot\phi_1 + 2ml^2\ddot\phi_2\cos(\phi_1-\phi_2) \\
        &\quad - 2ml^2\dot\phi_2\sin(\phi_1-\phi_2)(\dot\phi_1-\dot\phi_2) \\
        &\quad + ml^2\ddot\phi_3\cos(\phi_1-\phi_3) \\
        &\quad - ml^2\dot\phi_3\sin(\phi_1-\phi_3)(\dot\phi_1-\dot\phi_3) 
    \end{align}
    \begin{align}
        -3mgl\sin\phi_1 &= 3ml^2\ddot\phi_1 + 2ml^2\ddot\phi_2\cos(\phi_1-\phi_2)
            + ml^2\ddot\phi_3\cos(\phi_1-\phi_3)  \\
            &\quad + 2ml^2\dot\phi_2^2\sin(\phi_1-\phi_2)
            + ml^2\dot\phi_3^2\sin(\phi_1-\phi_3)
    \end{align}
    \begin{align}
        -\frac{3g}{l}\sin\phi_1 &= 3\ddot\phi_1 + 2\ddot\phi_2\cos(\phi_1-\phi_2)
            + \ddot\phi_3\cos(\phi_1-\phi_3) + 2\dot\phi_2^2\sin(\phi_1-\phi_2)
            + \dot\phi_3^2\sin(\phi_1-\phi_3)   \label{eq_m1}
    \end{align}

    For $\phi_2$ we have
    \begin{align}
        \pD{\L}{\phi_2} &= \D{}{t}\pD{\L}{\dot\phi_2} 
    \end{align}
    \begin{align}
        -2mgl\sin\phi_2 +2ml^2\dot\phi_1\dot\phi_2\sin(\phi_1-\phi_2)
            -ml^2\dot\phi_2\dot\phi_3\sin(\phi_2-\phi_3)
        &= 2ml^2\ddot\phi_2 + 2ml^2\ddot\phi_1\cos(\phi_1-\phi_2) \\
        &\quad - 2ml^2\dot\phi_1\sin(\phi_1-\phi_2)(\dot\phi_1-\dot\phi_2) \\
        &\quad + ml^2\ddot\phi_3\cos(\phi_2-\phi_3) \\
        &\quad - ml^2\dot\phi_3\sin(\phi_2-\phi_3)(\dot\phi_2-\dot\phi_3) 
    \end{align}
    \begin{align}
        -2mgl\sin\phi_2 &= 2ml^2\ddot\phi_2 + 2ml^2\ddot\phi_1\cos(\phi_1-\phi_2)
            + ml^2\ddot\phi_3\cos(\phi_2-\phi_3)  \\
            &\quad - 2ml^2\dot\phi_1^2\sin(\phi_1-\phi_2)
            + ml^2\dot\phi_3^2\sin(\phi_2-\phi_3)
    \end{align}
    \begin{align}
        -\frac{2g}{l}\sin\phi_2 &= 2\ddot\phi_2 + 2\ddot\phi_1\cos(\phi_1-\phi_2)
            + \ddot\phi_3\cos(\phi_2-\phi_3) - 2\dot\phi_1^2\sin(\phi_1-\phi_2)
            + \dot\phi_3^2\sin(\phi_2-\phi_3)   \label{eq_m2}
    \end{align}

    For $\phi_3$ we have
    \begin{align}
        \pD{\L}{\phi_3} &= \D{}{t}\pD{\L}{\dot\phi_3} 
    \end{align}
    \begin{align}
        -mgl\sin\phi_3 +ml^2\dot\phi_1\dot\phi_3\sin(\phi_1-\phi_3)
            +ml^2\dot\phi_2\dot\phi_3\sin(\phi_2-\phi_3)
        &= ml^2\ddot\phi_3 + ml^2\ddot\phi_1\cos(\phi_1-\phi_3) \\
        &\quad - ml^2\dot\phi_1\sin(\phi_1-\phi_3)(\dot\phi_1-\dot\phi_3) \\
        &\quad + ml^2\ddot\phi_2\cos(\phi_2-\phi_3) \\
        &\quad - ml^2\dot\phi_2\sin(\phi_2-\phi_3)(\dot\phi_2-\dot\phi_3) 
    \end{align}
    \begin{align}
        -mgl\sin\phi_3 &= ml^2\ddot\phi_3 + ml^2\ddot\phi_1\cos(\phi_1-\phi_3)
            + ml^2\ddot\phi_2\cos(\phi_2-\phi_3)  \\
            &\quad - ml^2\dot\phi_1^2\sin(\phi_1-\phi_3)
            - ml^2\dot\phi_2^2\sin(\phi_2-\phi_3)
    \end{align}
    \begin{align}
        -\frac{g}{l}\sin\phi_3 &= \ddot\phi_3 + \ddot\phi_1\cos(\phi_1-\phi_3)
            + \ddot\phi_2\cos(\phi_2-\phi_3) - \dot\phi_1^2\sin(\phi_1-\phi_3)
            - \dot\phi_2^2\sin(\phi_2-\phi_3)   \label{eq_m3}
    \end{align}

    Rearranging $\eqref{eq_m1}$, $\eqref{eq_m2}$, and $\eqref{eq_m3}$ by moving
    the second derivatives to one side, we have
    \begin{align}
         3\ddot\phi_1 + 2\ddot\phi_2\cos(\phi_1-\phi_2) + \ddot\phi_3\cos(\phi_1-\phi_3) + 
             &= -2\dot\phi_2^2\sin(\phi_1-\phi_2) - \dot\phi_3^2\sin(\phi_1-\phi_3)
             -\frac{3g}{l}\sin\phi_1 \\
        2\ddot\phi_1\cos(\phi_1-\phi_2)+ 2\ddot\phi_2 + \ddot\phi_3\cos(\phi_2-\phi_3)
            &= 2\dot\phi_1^2\sin(\phi_1-\phi_2) - \dot\phi_3^2\sin(\phi_2-\phi_3)
            -\frac{2g}{l}\sin\phi_2 \\
        \ddot\phi_1\cos(\phi_1-\phi_3)+\ddot\phi_2\cos(\phi_2-\phi_3)+\ddot\phi_3
            &= + \dot\phi_1^2\sin(\phi_1-\phi_3) + \dot\phi_2^2\sin(\phi_2-\phi_3)
            -\frac{g}{l}\sin\phi_3
    \end{align}
    We can rewrite this system  equations as a single matrix equation:
    \begingroup 
    \begin{equation} 
        \renewcommand*{\arraystretch}{1.75} % more vertical space in matrix
        \resizebox{0.95\textwidth}{!}{$
        \begin{pmatrix}
            3                    & 2\cos(\phi_1-\phi_2) & \cos(\phi_1-\phi_3) \\
            2\cos(\phi_1-\phi_2) & 2                    & \cos(\phi_2-\phi_3) \\
            \cos(\phi_1-\phi_3)  & \cos(\phi_2-\phi_3)  & 1                   \\
        \end{pmatrix}
        \begin{pmatrix}
            \ddot\phi_1 \\
            \ddot\phi_2 \\
            \ddot\phi_3
        \end{pmatrix} =
        \begin{pmatrix}
            -\dot\phi_3^2\sin(\phi_1-\phi_3) -2\dot\phi_2^2\sin(\phi_1-\phi_2)
                -\frac{3g}{l}\sin\phi_1 \\
            -\dot\phi_3^2\sin(\phi_2-\phi_3) + 2\dot\phi_1^2\sin(\phi_1-\phi_2)
                -\frac{2g}{l}\sin\phi_2 \\
            \dot\phi_2^2\sin(\phi_2-\phi_3) + \dot\phi_1^2\sin(\phi_1-\phi_3)
                -\frac{g}{l}\sin\phi_3
        \end{pmatrix}
        $}
    \end{equation} 
    \endgroup 
    We can then solve for our vector of second derivatives by multiplying on the
    left by the inverse of the first matrix:
    \begingroup 
    \begin{equation} \renewcommand*{\arraystretch}{1.75} % more vertical space in matrix
        \resizebox{0.95\textwidth}{!}{$
        \begin{pmatrix}
            \ddot\phi_1 \\
            \ddot\phi_2 \\
            \ddot\phi_3
        \end{pmatrix} =
        \begin{pmatrix}
            3                    & 2\cos(\phi_1-\phi_2) & \cos(\phi_1-\phi_3) \\
            2\cos(\phi_1-\phi_2) & 2                    & \cos(\phi_2-\phi_3) \\
            \cos(\phi_1-\phi_3)  & \cos(\phi_2-\phi_3)  & 1                   \\
        \end{pmatrix}^{-1}
        \begin{pmatrix}
            -\dot\phi_3^2\sin(\phi_1-\phi_3) -2\dot\phi_2^2\sin(\phi_1-\phi_2)
                -\frac{3g}{l}\sin\phi_1 \\
            -\dot\phi_3^2\sin(\phi_2-\phi_3) + 2\dot\phi_1^2\sin(\phi_1-\phi_2)
                -\frac{2g}{l}\sin\phi_2 \\
            \dot\phi_2^2\sin(\phi_2-\phi_3) + \dot\phi_1^2\sin(\phi_1-\phi_3)
                -\frac{g}{l}\sin\phi_3
        \end{pmatrix}
        $} \label{mat-eqn}
    \end{equation} 
    \endgroup 
    The matrix inverse in the above equation is very tedious to calculate. We
    used a computer algebra system\footnote{SageMath, the Sage Mathematics
        Software System (Version 7.6), The Sage Developers, 2017,
        \texttt{http://www.sagemath.org}.}
    to do the inverse and the matrix multiplication afterwards. The result is
    \begingroup 
    \begin{equation} 
        \renewcommand*{\arraystretch}{2.5} % more vertical space in matrix
        \resizebox{0.98\textwidth}{!}{$
        \begin{pmatrix}
            \ddot\phi_1 \\
            \ddot\phi_2 \\
            \ddot\phi_3
        \end{pmatrix} =
        \begin{pmatrix}
            \frac{1}{3} {\left(2 \dot\phi_2^2 \sin(\phi_1-\phi_2)-\dot\phi_3^2 \sin(\phi_1-\phi_3)-\frac{3 g \sin\left( \phi_1\right)}{l}\right)} {\left(\frac{2 {\left(\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)} \cos(\phi_1-\phi_2)}{2 \cos(\phi_1-\phi_2)^2- 3}-\cos(\phi_1-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6} + \frac{2 \cos(\phi_1-\phi_2)^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 1\right)}- {\left(2 \dot\phi_1^2 \sin(\phi_1-\phi_2)-\dot\phi_3^2 \sin(\phi_2-\phi_3)-\frac{2 g \sin\left(\phi_1\right)}{l}\right)} {\left(\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)} {\left(\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)} \cos(\phi_1-\phi_2)}{2 \cos(\phi_1-\phi_2)^2- 3}-\cos(\phi_1-\phi_3)\right)}}{{\left(2 \cos(\phi_1-\phi_2)^2- 3\right)} {\left(2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6\right)}} + \frac{\cos(\phi_1-\phi_2)}{2 \cos(\phi_1-\phi_2)^2- 3}\right)} + \frac{2 {\left(\dot\phi_1^2 \sin(\phi_1-\phi_3)-\dot\phi_2^2 \sin(\phi_2-\phi_3)-\frac{g \sin\left(\phi_1\right)}{l}\right)} {\left(\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)} \cos(\phi_1-\phi_2)}{2 \cos(\phi_1-\phi_2)^2- 3}-\cos(\phi_1-\phi_3)\right)}}{2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6} \\
            -{\left(2 \dot\phi_2^2 \sin(\phi_1-\phi_2)-\dot\phi_3^2 \sin(\phi_1-\phi_3)-\frac{3 g\sin\left(\phi_1\right)}{l}\right)} {\left(\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)} {\left(\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)} \cos(\phi_1-\phi_2)}{2 \cos(\phi_1-\phi_2)^2- 3}-\cos(\phi_1-\phi_3)\right)}}{{\left(2 \cos(\phi_1-\phi_2)^2- 3\right)} {\left(2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6\right)}} + \frac{\cos(\phi_1-\phi_2)}{2 \cos(\phi_1-\phi_2)^2- 3}\right)} + \frac{3}{2} {\left(2 \dot\phi_1^2 \sin(\phi_1-\phi_2)-\dot\phi_3^2 \sin(\phi_2-\phi_3)-\frac{2 g \sin\left(\phi_1\right)}{l}\right)} {\left(\frac{1}{2 \cos(\phi_1-\phi_2)^2- 3} + \frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{{\left(2 \cos(\phi_1-\phi_2)^2- 3\right)}^2 {\left(2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6\right)}}\right)}-\frac{3 {\left(\dot\phi_1^2 \sin(\phi_1-\phi_3)-\dot\phi_2^2 \sin(\phi_2-\phi_3)-\frac{g \sin\left(\phi_1\right)}{l}\right)} {\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}}{{\left(2 \cos(\phi_1-\phi_2)^2- 3\right)} {\left(2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6\right)}} \\
           \frac{2 {\left(2 \dot\phi_2^2 \sin(\phi_1-\phi_2)-\dot\phi_3^2 \sin(\phi_1-\phi_3)-\frac{3 g \sin\left(\phi_1\right)}{l}\right)} {\left(\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)} \cos(\phi_1-\phi_2)}{2 \cos(\phi_1-\phi_2)^2- 3}-\cos(\phi_1-\phi_3)\right)}}{2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6} + \frac{6 {\left(\dot\phi_1^2 \sin(\phi_1-\phi_3)-\dot\phi_2^2 \sin(\phi_2-\phi_3)-\frac{g \sin\left(\phi_1\right)}{l}\right)}}{2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6}-\frac{3 {\left(2 \dot\phi_1^2 \sin(\phi_1-\phi_2)-\dot\phi_3^2 \sin(\phi_2-\phi_3)-\frac{2 g \sin\left(\phi_1\right)}{l}\right)} {\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}}{{\left(2 \cos(\phi_1-\phi_2)^2- 3\right)} {\left(2 \cos(\phi_1-\phi_3)^2-\frac{{\left(2 \cos(\phi_1-\phi_2) \cos(\phi_1-\phi_3)- 3 \cos(\phi_2-\phi_3)\right)}^2}{2 \cos(\phi_1-\phi_2)^2- 3}- 6\right)}}
        \end{pmatrix}
        $}\, ,\vspace{0.5em}
    \end{equation} 
    \endgroup 
    which is far too messy to use computationally. Instead, we chose to
    construct the matrices in $\eqref{mat-eqn}$ in \texttt{numpy} and
    numerically calculate the inverse and multiplication. Since many of the
    terms are repeated (especially the trigonometric functions of angle
    differences), this has the added advantage of allowing us to precompute
    these functions.

    \end{appendices}

\end{document}